---
title: "data_visualization"
author: "Daniel Kerstan"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
#for ggplot
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)
library(sf)
```

## Get the lizards data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```
Histogram of lizards weights
```{r}
ggplot(lizards, aes(x = weight)) +
  geom_histogram(fill = "orange",
                 color = "red",
                 size = 1,
                 linetype = "dotted")
```

```{r}
# total_length on x-axis, weight on y-axis

ggplot(data = lizards, aes(x= total_length, y = weight)) + 
  geom_point(shape = 22,
             fill = "yellow",
             color = "red",
             size = 3,
             alpha = 0.5)
```
Bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length. total_length (x) , weight (y)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = common_name,
                 size = total_length),
             shape = 22,
             fill = "black"
             ) +
  theme_dark()
```

```{r}
ggplot(data = lizards, aes(x= total_length, y = weight)) + 
  geom_point(aes(color = common_name)) +
  facet_wrap(~common_name, ncol = 3)
```

```{r}
ggplot(data = lizards, aes(x= total_length, y = weight)) + 
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
  
```
Find total lizard counts by different groups in common name
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

#lizard_counts <- lizards %>% count(common_name)

#lizard_counts_cn_tail <- lizards %>% count(common_name, tail)

```

```{r}
ggplot(lizard_counts, aes(y = forcats::fct_reorder(common_name,count), x = count)) +
  geom_col(color = "red",
           fill = "black") +
  labs(x = "Number of lizards", y = "common name")
```
convert common_name to an ordered factor outside of ggplot
```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))
```


## Axis scale breaks, limits and labels
scatterplot: total_length vs weight (y)
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 50 , 500), 
                     limits =c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```
Transform the date column to class Date, then find counts of observations by date
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())
  
```

```{r}
ggplot(lizard_counts, aes(x = date, y = count)) + 
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%y")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple"),
                     breaks = c(0, 20, 40, 50))
```
### Update a color scheme using a palette in paletteer

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(lizards_fct, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name), 
               show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::default_gsea")
  
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "cyan4"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "white")
    
  ) +
  scale_x_continuous(breaks = c(100, 200))
```

